---
title: "HW 3 Key"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. (4 points)
\emph{From Lohr Exercise 2.10} Which of the following SRS designs will give the most precision (smallest variance) for estimating the population mean? Assume that each population has the same value of the population variance $S^2$. Please report the variance of the estimator for each scenario as a function of $S^2$ and then state which is the smallest.

1. An SRS of size 400 from a population of size 4000
2. An SRS of size 30 from a population of size 300
3. An SRS of size 3000 from a population of size 300,000,000

Recall the variance of an estimator of population mean is 
$$V(\hat{\overline{y_U}})=\frac{S^2}{n}\left(\frac{N-n}{N}\right)$$

1. $V(\hat{\overline{y_U}})=\frac{S^2}{400}\left(\frac{4000-400}{4000}\right) = \frac{S^2}{400} \times 0.9 = .00225 S^2$
\vfill
2. $V(\hat{\overline{y_U}})=\frac{S^2}{30}\left(\frac{300-30}{300}\right) = \frac{S^2}{30} \times 0.9 = .03 S^2$
\vfill
3. $V(\hat{\overline{y_U}})=\frac{S^2}{3000}\left(\frac{300,000,000-3000}{300,000,000}\right) = \frac{S^2}{3000} \times 0.999 =.000333 S^2$

So option 3 has the smallest variance.
\vfill

#### 2. 
Using the bird data set from Lab 2, a SRS sample of size 15 grid squares has been drawn for you. 
```{r}
birds <- read.csv('http://math.montana.edu/ahoegh/teaching/stat446/birdsurvey.csv', header = T)
set.seed(09252019)
N <- 200
n <- 15
sample_values <- sample(birds$bird.counts, size = n)
```

##### a. (4 points)
Compute an estimate of the _total_ number of birds across the entire region composed of 200 grid squares.

```{r}
t_hat <- mean(sample_values) * N
```

The estimate for the total number of birds is `r round(t_hat)`

##### b. (4 points)
Using the FPC compute the variance of the estimator for the population total.

```{r}
var_t <- (sd(sample_values)^2/n) * N * (N - n)
```

The variance for $\hat{t}$ is `r var_t`.

##### c. (4 points)
Construct a 95\% confidence interval for the population total.

```{r}
alpha <- .05
t_ci <- t_hat + c(-1,1) * qt(1-alpha/2, df = n-1) * sqrt(var_t)
```
The confidence interval for the population total is `r round(t_ci)`.


##### d. (4 points)
Discuss the similarities and differences for calculating uncertainty in our point estimates between this approach and the repeated sampling procedure from Lab 3.

_In the lab we took multiple samples to empirically construct the sampling distribution. In this setting, we only took a single samples. Hence, uncertainty is based on the difference in the sampling units through the sample standard deviation._


